### 2022-06-20 manuscript/manuscript.Rmd
update on table and ms

```{r}
# Sample
n_sample <- dataset[, .N]
#> 234

@@

From a total of `r n_sample` subjects with congenital hypotonia, `r n_sex["F", stat]` were females and `r n_sex["M", stat]` males (`r prop_sex`). The developmental characteristics of the sample can be seen in Table 1.

```{r}
mod <- gam(
  comunicacion_total ~ 
    s(profesional_id, sexo_paciente, respondedor_vinculo, bs = "re") + 
    s(edad_cronologica_meses), 
  data = dataset, 
  method = "REML"
)

gam.check(mod)

# Overall slope
(gam_coeffs <- report_overall(mod, "edad_cronologica_meses"))
(gam_slopes <- report_slopes(mod, "edad_cronologica_meses"))
(gam_smooth <- report_smooth(mod))

plot_gam(dataset, edad_cronologica_meses, comunicacion_total, round_to = 5,
         "Chronological age (months)", "Communication score", method = "REML")
```

When modelling the effect of chronological age on developmental skills, we observed a significant non-linear relationship on communication scores (`r gam_smooth[1, expr]`), that reflect an overall negative marginal effect (`r gam_coeffs`), however, this was not true when assessing the direction of the effect in the age range between 1 to 7 (`r gam_slopes[1, expr]`), and neither in the 18 to 48 months old group (`r gam_slopes[3, expr]`), whereas the effect tend to be positive but non-significant. The relationship between developmental skills and corrected age can be seen in Figure 1.

```{r eval=FALSE}
mod <- gam(
  comunicacion_total ~ 
    s(profesional_id, respondedor_vinculo, sexo_paciente, bs = "re") + 
    s(edad_corregida_meses), 
  data = dataset, 
  method = "REML"
)

gam.check(mod)

# Overall slope
(gam_coeffs <- report_overall(mod, "edad_corregida_meses"))
(gam_slopes <- report_slopes(mod, "edad_corregida_meses"))
(gam_smooth <- report_smooth(mod))

plot_gam(dataset, edad_corregida_meses, comunicacion_total, round_to = 5,
         "Corrected age (months)", "Communication score", method = "REML")
```

@@

# Data Availability Statement

@@

**Figure 1**. Relationship between corrected age (in months) and developmental domains. Regression lines represent fitted values estimated from GAM models. Points and error bars represent the mean and standard error at 5-month age intervals.

>>>>

```{r eval=FALSE}
# Sample
n_sample <- dataset[, .N]
#> 234

@@

From a total of 234 subjects with congenital hypotonia, 94 (40.2%) were females and 140 (59.8%) males ($\chi^2_{gof}$ (1) = 9.04, *p* = 0.003, $C_{Pearson}$ = 0.19, CI~0.95%~[0.09, 1]). The developmental characteristics of the sample can be seen in [Table 1](#tab1).

```{r eval=FALSE}
mod <- gam(
  comunicacion_total ~ 
    s(profesional_id, sexo_paciente, respondedor_vinculo, bs = "re") + 
    s(edad_corregida_meses), 
  data = dataset, 
  method = "REML"
)

report_overall(mod, "edad_corregida_meses")
#> [1] "$\\beta$ = -2.36, CI~95%~[-3.47, -1.25], $t_{student}$ (224.04) = -4.2, *p* < 0.001"

report_slopes(mod, "edad_corregida_meses")
#>    Start  End edad_corregida_meses                                                                                 expr
#> 1:   0.0  6.8                  3.4    $\\beta$ = 0.49, CI~95%~[-0.89, 1.86], $t_{student}$ (224.04) = 0.45, *p* = 0.319
#> 2:   7.3 17.9                 12.6 $\\beta$ = -2.08, CI~95%~[-3.17, -0.99], $t_{student}$ (224.04) = -3.79, *p* = 0.004
#> 3:  18.4 48.0                 33.2    $\\beta$ = 0.45, CI~95%~[-1.32, 2.23], $t_{student}$ (224.04) = 0.42, *p* = 0.593

report_smooth(mod)
#>                                                   term                                            expr
#> 1:                             s(edad_corregida_meses) $F_{smooth}$ (5.2, 224.04) = 13.43, *p* < 0.001
#> 2: s(profesional_id,sexo_paciente,respondedor_vinculo)  $F_{smooth}$ (3.76, 224.04) = 0.72, *p* = 0.02
```

When modelling the effect of chronological age corrected for prematurity on developmental domains, we observed a significant non-linear relationship on communication scores ($F_{smooth}$ (5.2, 224.04) = 13.43, *p* < 0.001), that reflect an overall negative marginal effect ($\beta$ = -2.36, CI~95%~[-3.47, -1.25], $t_{student}$ (224.04) = -4.2, *p* < 0.001), however, this was not true when assessing the direction of the effect in the age range between 1 to 7 ($\beta$ = 0.49, CI~95%~[-0.89, 1.86], $t_{student}$ (224.04) = 0.45, *p* = 0.319), and neither in the 18 to 48 months old group ($\beta$ = 0.45, CI~95%~[-1.32, 2.23], $t_{student}$ (224.04) = 0.42, *p* = 0.593), whereas the effect tend to be positive but non-significant.

```{r eval=FALSE}
mod <- gam(
  motora_gruesa_total ~ 
    s(profesional_id, sexo_paciente, respondedor_vinculo, bs = "re") + 
    s(edad_corregida_meses), 
  data = dataset, 
  method = "REML"
)

report_overall(mod, "edad_corregida_meses")
#> [1] "$\\beta$ = 1.95, CI~95%~[0.66, 3.25], $t_{student}$ (226.75) = 2.97, *p* = 0.003"

report_slopes(mod, "edad_corregida_meses")
#>    Start  End edad_corregida_meses                                                                                expr
#> 1:   0.0  6.8                  3.4 $\\beta$ = -2.94, CI~95%~[-4.55, -1.34], $t_{student}$ (226.75) = -3.7, *p* = 0.004
#> 2:   7.3  9.2                  8.2    $\\beta$ = 0.02, CI~95%~[-1.12, 1.17], $t_{student}$ (226.75) = 0.06, *p* = 0.45
#> 3:   9.7 15.5                 12.6    $\\beta$ = 1.86, CI~95%~[0.61, 3.11], $t_{student}$ (226.75) = 2.93, *p* = 0.009
#> 4:  16.0 48.0                 32.0  $\\beta$ = -0.02, CI~95%~[-2.05, 2.01], $t_{student}$ (226.75) = 0.07, *p* = 0.646

report_smooth(mod)
#>                                                   term                                            expr
#> 1:                             s(edad_corregida_meses) $F_{smooth}$ (5.24, 226.75) = 6.19, *p* < 0.001
#> 2: s(profesional_id,sexo_paciente,respondedor_vinculo)  $F_{smooth}$ (1.01, 226.75) = 0.1, *p* = 0.244
```

When analyzing the motor skills domain we found a significant non-linear effect of corrected age on gross motor scores ($F_{smooth}$ (5.24, 226.75) = 6.19, *p* < 0.001), which had an overall positive effect ($\beta$ = 1.95, CI~95%~[0.66, 3.25], $t_{student}$ (226.75) = 2.97, *p* = 0.003), nevertheless, the slope varied across age, whereas between the 0 to 7 age range this has a negative effect on gross motor score $\beta$ = -2.94, CI~95%~[-4.55, -1.34], $t_{student}$ (226.75) = -3.7, *p* = 0.004, but in the 9.7 to 15.5 a positive effect was observed ($\beta$ = 1.86, CI~95%~[0.61, 3.11], $t_{student}$ (226.75) = 2.93, *p* = 0.009), meanwhile in the 7.3 to 9.2 ($\beta$ = 0.02, CI~95%~[-1.12, 1.17]) and 16 to 48 months old ($\beta$ = -0.02, CI~95%~[-2.05, 2.01]) the slope was non significant and practically cero ($t_{student}$ (226.75) = 0.06, *p* = 0.45, and $t_{student}$ (226.75) = 0.07, *p* = 0.646 respectively).

The relationship between developmental domains and corrected age can be seen in [Figure 1](#fig1).

@@

@@

[**Table 1**]{#tab1}. Overall baseline and developmental characteristics of the sample and grouped by sex. ^1^ Data is presented as sample size, and *Mdn* (*IQR*); ^2^ P-values derived from the *Wilcoxon* rank-sum test.

[**Figure 1**]{#fig1}. Relationship between corrected age (in months) and developmental domains. Regression lines and shaded area represent predicted values estimated from GAM models and their 95% confidence interval. Points and error bars represent the mean and standard error at 5-month age intervals.
